Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-10-22 papers/thesis-kirstie/thesis.tex
updates and figures

\abstract{
  This is a test
}

@@

Classically, fluids have been modeled by a collection of small, impenetrable spheres colliding 
with one another like billiard balls. In the simplest case, the spheres are treated as having 
only a replulsive potential. This model, called the Hard-Sphere fluid, has served as the primary 
reference system for fluids. This is due to the fact that the repulsion between atoms is much 
stronger than their attraction, and so the repulsive potential, which is descriptive of 
hard spheres, dominates the interactions between atoms. 
In 1954 Zwanzig introduced the Square-well fluid which is a perturbation of the Hard-Sphere fluid 
created by adding a hard-edge attractive potential to the repulsive potential of the Hard-Sphere 
fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied spheres with softly fading replusive potentials which he 
modeled by making the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
This reflects the idea that soft-spheres penetrate more deeply into each other 
when they collide at higher temperatures. 
%At higher temperatures the spheres would penetrate more deeply into each other when they collide, 
%and thus be represented by spheres with smaller radii. 
In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
with Rowlinson's idea of using temperature-dependent radii to model spheres with 
soft repulsive potentials to create the Barker-Henderson model fluid~\cite{barker1967perturbation}, 
where a soft-sphere fluid by is modeled by using a Hard-Sphere fluid with hard spheres of 
larger radii at low temperatures, and smaller radii at high temperatures. 
%
%an expression for the excess Helmholtz free energy of inhomogeneous fluids.
% ?  Percus-Yevick applied Density Functional Theory (DFT) to find an expression for the excess 
%Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. 
%
%In 1967 Barker and Henderson~\cite{barker1967perturbation} adapted the perturbation method of Zwanzig 
%to use with fluids composed of spheres with softly fading potentials rather than the hard edges of the spheres 
%in the Hard-Sphere fluid. 
%This was done by making the radius of the spheres temperature dependent, 
%so that at low temperatures the Soft-Sphere fluid could be modeled by hard spheres with large radii, 
%and at high temperatures the Soft-Sphere fluid could be modeled by hard spheres with smaller radii. 
%
%A larger radii cooresponds to a deeper level of penetration of two soft spheres into each other 
%when they collide at higher temperatures.
%The different size radii coorespond to the different depths to which two soft 
%spheres would penetrate into each other when they collide. 
%This reflects the idea that soft-spheres penetrate more deeply into each other when 
%they collide at higher temperatures. 
In 1971 Weeks, Chandler, and Anderson used temperature and density dependent hard-sphere 
radii to model spheres with the more realistic softly fading repulsive portion of the 
Leonard-Jones potential~\cite{andersen1971relationship}, which is known as the 
Weeks-Chandler-Anderson model fluid.

While models for particular fluids were being created, powerful methods to describe the properties 
of fluids in general were also being developed. In the late 1950s, Scaled Particle 
Theory (SPT)~\cite{ReissSPT} was introduced from which an equation of state, and
an expression for the excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
obtained~\cite{Hansen}. \color{red}ADD Percus-Yevick introduced an alternative method...OZ... \color{black}
%Percus and Yevick took another approach to the Hard-Sphere fluid.
In 1961 Morita and Hiroike introduced a density functional approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
(the equilibrium Helmholtz free energy in the case of a canonical ensemble), and the cooresponding equilibrium 
number density profile for inhomogeneous fluids. To use this approach, a functional must be 
constructed for a particular fluid.
%DFT Evans 1979
Led by the consistency bewteen results from SPT and Percus-Yevick, in 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
(which can be constructed from SPT~\cite{santos2012phi3}) to form a functional for 
%which provides an expression for the excess Helmholtz free energy of an inhomogeneous 
the Hard-Sphere fluid. 
Tarazona later improved on the method by introducing a tensor version of the 
functional~\cite{tarazonaphi3, santos2012phi3}, and in 1999 Schmidt adopted FMT for soft spheres 
forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.

Schmidt applied his Soft Fundamental Measure Theory to various fluids with soft repulsive 
and soft attractive potentials, and developed an erf model for 
spheres with steep, short-range soft repulsive potentials~\cite{schmidt2000fluid}. In this paper, 
Density Functional Theory (DFT) will be used to predict the freezing of a 
Weeks-Chandler-Anderson model fluid by applying SFMT with Schmidt's proposed 
erf model to form the functional for the excess Helmholtz free energy. 
%for generating the weight funcions, and Tarazona's tensor functional. 
Results will be compared to those generated from Monte-Carlo simulations.

@@

Real fluids and their behavior are very complex from a physicist's point of view. Although a water molecule, for example, is composed of only three atoms, the ever-changing redistribution of charges within the molecules, and their interaction with charges from other molecules make it difficult to study a fluid composed of water molecules directly. The task becomes even more daunting considering the complicated molecular motion of each water molecule, and the sheer number of molecules that must be taken into account. 
%Although a water molecule, for example, is composed of only three atoms, the ever-changing redistribution of charges within the molecules and their interaction with charges from other molecules, the complicated molecular motion of a collection of water molecules, and the sheer number of molecules that must be taken into account make it difficult to study a fluid composed of water molecules directly.
The problem of understanding the behavoir of real liquids is better addressed by using models which reduce a complex picture to a simple one. Modifications, or perturbations, can then be incorporated into the model making it progressively more complex until it mimics the behavior of a real liquid as closely as possible.

@@

Another fairly simple, but relatively accurate interatomic potential widely used to represent generic, neutral atoms is the Leonard-Jones potential. The Leonard-Jones potential describes both the long-range attraction, and the short-range replusion that occurs between two nuetral atoms. The attraction between two nuetral atoms is due to fleeting dipoles that form spontaneously when electrons are redistributed as atoms come into close proximty. These are often characterized as Van der Walls, or London dispersion forces. The attraction grows stronger as two atoms move closer together until their atomic orbitals begin to overlap. %At this point the electrons in each atom begin to repel the electrons in the other atom and oppose the overlap of the two atoms.
At this point the electrons in each atom begin to oppose the overlap of the two atoms through their replusion. The repulsion increases rapidly and approaches infinity as the distance between the atoms conintues to decrease.

@@

The repulsive portion of the Leonard-Jones potential forms the potential for a Weeks-Chandler-Anderson (WCA) model fluid, shown in Figure \ref{fig:WCA_potential}. The WCA potential is described by the equation: \begin{equation}{V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.}\end{equation}

\noindent As seen in Figure \ref{fig:WCA_potential}, the WCA potential decreases from infinity as the distance between the centers of the atoms r increases, and becomes, and remains, essentially zero when r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ where R is the radius of the sphere as shown in Figure \ref{fig:TwoSpheres}. As with the Lenard-Jones potential, the repulsive WCA potential increases rapidly and approches infinity as the two spheres begin to overlap. This is a more realistic repulsive potential than that of the Hard Sphere model which goes instantly to infinity when the two spheres make contact (at r=2R), and is zero otherwise. Because the WCA spheres have no hard boundary in their interatomic potential, they are called ``soft spheres'' and their gradually fading potential makes them slightly ``squishy''.

@@

The WCA potential is not well suited for implementing Density Functional Theory.  
Therefore, an erf potential which closely approximates the WCA potential at temperatures 
of interest will be used instead of the WCA potential. The erf potential is given by 
\begin{equation}
	{V_{erf}=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]}
\end{equation} 
where the parameter $\alpha$ is derived by setting the erf potential equal to the WCA 
potentail when r equals $\alpha$ 
\begin{displaymath}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{displaymath} 
and is given by 
\begin{equation}\label{alphaT}
	{\alpha=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}}
\end{equation}

%Deleted old Xi method
%move 3-9 to methods

Unlike the WCA potential, the erf potential depends on temperature through the parameters $\alpha$ and $\Xi$. Figure \ref{fig:alphaXivsT} shows how $\alpha$ and $\Xi$ vary with temperature when $\sigma=1$ and $\epsilon=1$. % and Figure \ref{fig:erf_Potential} shows the erf potential at reduced temperatures $\frac{k_BT}{\epsilon}=1$, $\frac{k_BT}{\epsilon}=2$, and $\frac{k_BT}{\epsilon}=3$ alongside the WCA potential. %As seen in Figure \ref{fig:erf_potential}, the erf potentials closely match the WCA potential for reduced temperatures below $\frac{k_BT}{\epsilon}=2$, and for r/$\sigma$ less than ???. % so only values of $V_{erf}$ that are much less than $k_BT$/$\epsilon$ ??? will be considered. 
% In these regions $\alpha$ decreases with increasing temperature, and $\Xi$ increases with increasing temperature.

 \begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.4\linewidth}
      \includegraphics[height=5cm]{plot_alpha.pdf}
      \caption{$\alpha$ versus $k_BT/\epsilon$}
    \end{subfigure} 
    \begin{subfigure}[b]{0.4\linewidth}
      \includegraphics[height=5cm]{plot_xi.pdf}
      \caption{$\Xi$ versus $k_BT/\epsilon$}
    \end{subfigure} 
    \caption{These plots, generated with $\sigma=1$ and $\epsilon=1$, show how the parameters $\alpha$ and $\Xi$ vary with temperature. (a) $\alpha$ decreases with increasing temperature. (b) $\Xi$ rapidly increases for temperatures below $\frac{k_BT}{\epsilon}=52$ and then gradually decreases.}
    \label{fig:alphaXivsT}
    \end{figure}

%\begin{figure}[h!]
%    \centering
    %\includegraphics[height=7cm]{erf_Potential.pdf}
%    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ shown in black.}
%    \label{fig:erf_Potential}
%  \end{figure}

%\color{red}\noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!\color{black}

Another way to find values for $\Xi(T)$ is to use the second virial coefficent $B_{2}$ from the equation of state expressed in terms of the virial expansion. 
\begin{equation}\label{virialequation}
    \frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) 
\end{equation}
where 
\begin{equation}
	B_{2}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}(r)}-1\right)d\vec{r} 
\end{equation}
The second virial coefficient can be derived from Euler's equation for the internal energy of a system, 
\begin{equation}U=TS-PV+\mu{N}\end{equation}
using the relations from statistical thermophysics which views the thermodynamic quantities as ensemble averages
\begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
\begin{equation}U=<E>=\sum_{i=0}^\infty{P_iE_i}\end{equation}
\begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
which, for a canonical ensemble, gives
\begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
Expressing the Helmholtz free energy in terms of the pair potential $V(r)$ results in the equation of state expressed in terms of the virial expansion as given by Equation~\ref{virialequation}.

The value of $\Xi$ at a given temperature, T, can be found computationally by equating the second virial coefficient evaluated for the error function potential $V_{erf}(r)$, and the second virial coefficent evalued for the WCA potential $V_{WCA}$.
\begin{equation}B_{2erf}(T) =B_{2WCA}(T)\end{equation}
Evaluating $B_{2}$ for the erf potential gives
%\begin{equation}B_{2erf}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{erf}(r)}-1\right)d\vec{r} \end{equation}
\begin{equation}
	B_{2erf}(T) = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{erf}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]
\end{equation}
where $\alpha(T)$ is given by Equation~\ref{alphaT} as before. Similarly, $B_{2WCA}$ can be found computationally from 
\begin{equation}
	B_{2WCA}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{WCA}(r)}-1\right)d\vec{r} 
\end{equation}

A plot of $\Xi(T)$ obtained through this method is shown in Figure \ref{fig:xi_fromB2vsT}, and Figure \ref{fig:erf_Potential_xifromB2} shows the erf potential at a few different temperatures alongside the WCA potential.

\begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{xi_fromB2.pdf}
    \caption{Shows the temperature dependence of $\Xi$ generated from matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ at each temperature.}
    \label{fig:xi_fromB2vsT}
  \end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{erf_Potential_xifromB2.pdf}
    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ shown in black.}
    \label{fig:erf_Potential_xifromB2}
  \end{figure}

@@

Freezing is governed by entropy, which the universe, or any closed system, seeks to maximize according to the second law of Thermodynamics. When the entropy of a pure substance in its liquid state is lower than the entropy that it would have in its solid state at the same temperature (necessary for thermal equilibrium), the same pressure (necessary for mechanical equilibrium), and the same chemical potential (necessary for diffusive equilibrium), an abrupt phase transition occurs, that is, the liquid freezes forming a crystalline structure. The particular crystalline structure and density that result are affected by the interatomic potential. When repulsive, soft sphere fluids (that are not too soft) freeze, they form lattice structures composed of Face Centered Cubic (FCC) crystal cells, as do hard spheres. %This structure is the three-dimensional equivilant to the two-dimensional structure that one would see by filling the bottom of a box with a single layer of marbles. 
This structure is what one would see by filling a cube with marbles. It results in the highest density achievable for hard, and nearly-hard, spheres that are purely repulsive.

@@

%For each particular temperature, the pressure of the liquid and the solid are the same at the point of transistion. 
\begin{figure}[h!]
    \centering
    \includegraphics[height=5cm]{P-T_Diagram.png}
    \caption{Conceptual Pressure-Temperature Phase Diagram}
    \label{fig:P-T_Diagram}
  \end{figure}
The pressure can be related to the Helmholtz free energy F, which is minimized for a system at fixed temperature, volume and number of particles as the entropy of the system and its surroundings is maximized. The negative of the partial derivative of the Helmholtz free energy with respect to volume V at a fixed temperature T and number of particles N equals the pressure. 
\begin{equation}{P=-\frac{\partial{F}}{\partial{V}}\bigg|_{T,N}}\end{equation}
\noindent Dividing both the Helmholtz free energy and the volume by the number of particles, the partial derivative of the Helmholtz free energy with respect to volume becomes the partial derivative of the Helmholtz free energy per atom with respect to the inverse number density. \begin{equation}{P=-\frac{\partial{\frac{F}{N}}}{\partial{\frac{V}{N}}}\bigg|_{T,N} = -\frac{\partial{f}}{\partial{\frac{1}{n}}}\bigg|_{T,N}}\end{equation} Thus, the pressure corresponds to the slope of the curve on a plot of the Helmholtz free energy per atom versus the inverse density at fixed temperature.

%In a process called Maxwell's Double Tangent construction, two curves are used to find the densities of the liquid and the solid at the point of transistion. Figure \ref{fig:MaxwellDT} shows a curve for the liquid, and a curve for the solid both at a fixed temperature. Since the pressure of the liquid is the same as the pressure of the solid at the point of transistion, the density of the liquid and the density of the solid at the point of freezing can be found from identifying the inverse densties at which the tangents to each curve have the same slope.%
Two such curves are shown in Figure \ref{fig:MaxwellDT}, one for the liquid and one for the solid at a fixed temperature. Since the pressure of the liquid is the same as the pressure of the solid at the point of transistion, the density of the liquid and the density of the solid at the point of freezing can be found from identifying the inverse densties at which the tangents to each curve have the same slope. This process is called Maxwell's Double Tangent Construction.

\begin{figure}[h!]
    \centering
     \includegraphics[height=7cm]{figs/MaxwellDTC-Fig1.pdf}
     % FIXME\includegraphics[height=7cm]{plot-pressure_Fig1.pdf}
    \caption{Identifying densities at the solid-liquid transition using Maxwell's Double Tangent Construction}
    \label{fig:MaxwellDT}
  \end{figure}

@@

%\begin{displaymath}{F=-PV+G}\end{displaymath} 
\begin{displaymath}{\frac{F}{N}=-P\frac{V}{N}+\frac{G}{N}}\end{displaymath} 
\begin{displaymath}{f=-P\frac{1}{n}+g}\end{displaymath}
\noindent This equation has the form of a line:
\begin{displaymath}{y=mx+y_o}\end{displaymath}
%\begin{displaymath}{G_1=G_2}\end{displaymath} 
%\begin{displaymath}{F_1+PV_1=F_2+PV_2}\end{displaymath} 
%\begin{displaymath}{\frac{F_1}{N}+P\frac{V_1}{N}=\frac{F_2}{N}+P\frac{V_2}{N}}\end{displaymath} 
\noindent and describes the red line shown in Figure \ref{fig:MaxwellDT} with the pressure given by the negative of the slope of the line, and the Gibb's free energy per atom given by the y-intercept. At the point of phase transition:
\begin{displaymath}{g_L=g_S}\end{displaymath} 
\begin{displaymath}{f_L+P\frac{1}{n_L}=f_S+P\frac{1}{n_S}}\end{displaymath}

The pressure at the transistion point can be obtained from the point of intersection that occurs when the Gibbs free energy per atom is plotted against the pressure for both the liquid and the solid at fixed temperature, as shown in Figure \ref{fig:GibbsvsP}. These curves are constructed from the slope and cooresponding y-intercept data collected from lines tangent along the curves on the the Helmholtz free energy per atom verses inverse density plots for both the liquid and the solid. 
\begin{figure}[h!]
    \centering
     \includegraphics[height=7cm]{MaxwellDTC-Fig2.pdf}
     %\includegraphics[height=7cm]{plot-pressure_Fig2.pdf}
    \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
    \label{fig:GibbsvsP}
  \end{figure}
%CURVE FITTING! - put in methods/analysis

@@

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{MaxwellDTC-Fig3.pdf}
    % \includegraphics[height=7cm]{plot-pressure_Fig3.pdf}
    \caption{The liquid and solid densities at the point of transistion are found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
    \label{fig:Pvsinvn}
  \end{figure}

A Temperature-Density (T-n) phase diagram, such as that in Figure \ref{fig:T-n_Diagram}, can be generated by repeating this procedure for various temperatures. %The lowest density the crystal exhibts and the highest density the fluid exhibits at the phase transistion can then be identified for each temperature.

\begin{figure}[h!]
    \centering
    \includegraphics[height=4cm]{T-n_Diagram.png}
    \caption{Conceptual Temperature-Density Phase Diagram used to identify the number densities of the liquid and the solid during a phase transistion. The gray region is where the liquid and the solid coexist, and are therefore not pure states.}
    \label{fig:T-n_Diagram}
  \end{figure} 
%One way of implementing this practically is to determine what the Helmholtz free energy would be for both the liquid and solid phases at a particular temperature and density. Whichever phase has the lower Helmholtz free energy will be the phase that the substance will be in.

@@

%The reason this works is because cDFT satisfies the conditions of the variational principle, the same mathematical principle used to determine that the shortest distance between two points is a straight line by varying the path between the two points until the distance is minimized.

>>>>

\abstract{
  We use Classical Density Functional Theory (cDFT) to predict the 
  freezing of a Weeks-Chandler-Anderson (WCA) fluid. A WCA fluid 
  consists of soft spheres that exhibit the repulsive portion of the 
  Leonard-Jones potential widely used to model the interaction of two 
  neutral atoms. Our cDFT functional is constructed using Soft Fundamental 
  Theory (SFMT) with Tarazona's tensor expression for Phi3, and Schmidt's 
  erf model in which we have incorporated our temperature dependent 
  parameters to fit the results to those predicted by Monte-Carlo simulations. 
  Our functional was successful in predicting the freezing of the WCA fluid 
  with a fairly good match to Monte-Carlo simulations. This result suggests 
  that SFMT may provide a solid foundation for future density functionals 
  that will model real fluids.
}

@@

Classically, fluids have been modeled by a collection of small, 
impenetrable spheres colliding with one another like billiard balls. 
In the simplest case, the spheres are treated as having only a replulsive 
potential. This model, called the Hard-Sphere fluid, has served as the 
primary reference system for fluids. This is due to the fact that the 
repulsion between atoms is much stronger than their attraction, and so 
the repulsive potential, which is descriptive of hard spheres, dominates 
the interactions between atoms. In 1954 Zwanzig introduced the Square-well 
fluid which is a perturbation of the Hard-Sphere fluid created by adding 
a hard-edge attractive potential to the repulsive potential of the 
Hard-Sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
spheres with softly fading replusive potentials which he modeled by making
the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
This reflects the idea that soft-spheres penetrate more deeply into 
each other when they collide at higher temperatures.

In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
with Rowlinson's idea of using temperature-dependent radii to model spheres 
with soft repulsive potentials to create the Barker-Henderson model 
fluid~\cite{barker1967perturbation}, where a soft-sphere fluid by is 
modeled by using a Hard-Sphere fluid with hard spheres of larger radii 
at low temperatures, and smaller radii at high temperatures. In 1971 Weeks, 
Chandler, and Anderson used temperature and density dependent hard-sphere 
radii to model spheres with the more realistic softly fading repulsive 
portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
which is known as the Weeks-Chandler-Anderson model fluid.

While models for particular fluids were being created, powerful methods 
to describe the properties of fluids in general were also being developed. 
In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
introduced from which an equation of state, and an expression for the 
excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
obtained~\cite{Hansen}. \color{red}ADD Percus-Yevick introduced an 
alternative method...OZ... \color{black}
%Percus and Yevick took another approach to the Hard-Sphere fluid.
In 1961 Morita and Hiroike introduced a density functional 
approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
(the equilibrium Helmholtz free energy in the case of a canonical 
ensemble), and the cooresponding equilibrium number density profile for 
inhomogeneous fluids. To use this approach, a functional must be 
constructed for a particular fluid.
%DFT Evans 1979
%Led by the consistency bewteen results from SPT and Percus-Yevick, 
In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
(which can be constructed from SPT~\cite{santos2012phi3}) to form a 
functional for the Hard-Sphere fluid. Tarazona later improved on the 
method by introducing a tensor version of the functional~\cite{tarazonaphi3, 
santos2012phi3}, and in 1999 Schmidt adopted FMT for soft spheres 
forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.

Schmidt applied his Soft Fundamental Measure Theory to various fluids 
with soft repulsive and soft attractive potentials, and developed an erf 
model for spheres with steep, short-range soft repulsive 
potentials~\cite{schmidt2000fluid}. In this paper, Density Functional 
Theory (DFT) will be used to predict the freezing of a Weeks-Chandler-Anderson 
model fluid by applying SFMT with Schmidt's proposed erf model, and our temperature
dependent parameters to form the functional for the excess Helmholtz free energy. 
%for generating the weight funcions, and Tarazona's tensor functional. 
Results will be compared to those generated from Monte-Carlo simulations.

@@

Real fluids and their behavior are very complex from a physicist's point 
of view. Although a water molecule, for example, is composed of only three
atoms, the ever-changing redistribution of charges within the molecules, 
and their interaction with charges from other molecules make it difficult 
to study a fluid composed of water molecules directly. The task becomes 
even more daunting considering the complicated molecular motion of each 
water molecule, and the sheer number of molecules that must be taken 
into account.

The problem of understanding the behavoir of real liquids is better addressed 
by using models which reduce a complex picture to a simple one. Modifications, 
or perturbations, can then be incorporated into the model making it 
progressively more complex until it mimics the behavior of a real 
liquid as closely as possible.

@@

Another fairly simple, but relatively accurate interatomic potential widely 
used to represent generic, neutral atoms is the Leonard-Jones potential. 
The Leonard-Jones potential describes both the long-range attraction, 
and the short-range replusion that occurs between two nuetral atoms. 
The attraction between two nuetral atoms is due to fleeting dipoles that 
form spontaneously when electrons are redistributed as atoms come into 
close proximty. These are often characterized as Van der Walls, or London 
dispersion forces. The attraction grows stronger as two atoms move closer 
together until their atomic orbitals begin to overlap. 
%At this point the electrons in each atom begin to repel the electrons 
%in the other atom and oppose the overlap of the two atoms.
At this point the electrons in each atom begin to oppose the overlap of 
the two atoms through their replusion. The repulsion increases rapidly and 
approaches infinity as the distance between the atoms conintues to decrease.

@@

The repulsive portion of the Leonard-Jones potential forms the potential for a 
Weeks-Chandler-Anderson (WCA) model fluid, shown in Figure \ref{fig:WCA_potential}. 
The WCA potential is described by the equation: 
\begin{equation}{V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.}\end{equation} 
As seen in Figure \ref{fig:WCA_potential}, the WCA potential decreases 
from infinity as the distance between the centers of the atoms r increases, 
and becomes, and remains, essentially zero when r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ 
where R is the radius of the sphere as shown in Figure \ref{fig:TwoSpheres}. 
As with the Lenard-Jones potential, the repulsive WCA potential increases 
rapidly and approches infinity as the two spheres begin to overlap. This 
is a more realistic repulsive potential than that of the Hard Sphere model 
which goes instantly to infinity when the two spheres make contact (at r=2R), 
and is zero otherwise. Because the WCA spheres have no hard boundary in 
their interatomic potential, they are called ``soft spheres'' and their 
gradually fading potential makes them slightly ``squishy''.

@@

%MOVE to METHODS start >>>
%........................
%MOVE to METHODS  end  >>>

@@

Freezing is governed by entropy, which the universe, or any closed system, 
seeks to maximize according to the second law of Thermodynamics. When the 
entropy of a pure substance in its liquid state is lower than the entropy 
that it would have in its solid state at the same temperature (necessary 
for thermal equilibrium), the same pressure (necessary for mechanical 
equilibrium), and the same chemical potential (necessary for diffusive 
equilibrium), an abrupt phase transition occurs, that is, the liquid freezes 
forming a crystalline structure. The particular crystalline structure and 
density that result are affected by the interatomic potential. When repulsive, 
soft sphere fluids (that are not too soft) freeze, they form lattice 
structures composed of Face Centered Cubic (FCC) crystal cells~\cite{Hansen}, 
as do hard spheres. 
%This structure is the three-dimensional equivilant to the two-dimensional 
%structure that one would see by filling the bottom of a box with a single layer of marbles. 
This structure is what one would see by filling a cube with marbles. It 
results in the highest density achievable for hard, and nearly-hard, spheres 
that are purely repulsive.

@@

%For each particular temperature, the pressure of the liquid and the 
%solid are the same at the point of transistion. 
\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{P-T_Diagram.png}
    \caption{Conceptual Pressure-Temperature Phase Diagram}
    \label{fig:P-T_Diagram}
  \end{figure}
The pressure can be related to the Helmholtz free energy F, which is minimized 
for a system at fixed temperature, volume and number of particles as the 
entropy of the system and its surroundings is maximized. The negative of 
the partial derivative of the Helmholtz free energy with respect to volume V 
at a fixed temperature T and number of particles N equals the pressure. 
\begin{equation}{P=-\frac{\partial{F}}{\partial{V}}\bigg|_{T,N}}\end{equation}
\noindent Dividing both the Helmholtz free energy and the volume by the 
number of particles, the partial derivative of the Helmholtz free energy 
with respect to volume becomes the partial derivative of the Helmholtz 
free energy per atom with respect to the inverse number density. 
\begin{equation}{P=-\frac{\partial{\frac{F}{N}}}{\partial{\frac{V}{N}}}\bigg|_{T,N} = -\frac{\partial{f}}{\partial{\frac{1}{n}}}\bigg|_{T,N}}\end{equation} Thus, the pressure corresponds to the slope of the curve on a plot of the Helmholtz free energy per atom versus the inverse density at fixed temperature.

%In a process called Maxwell's Double Tangent construction, two curves are 
%used to find the densities of the liquid and the solid at the point of transistion. 
%Figure \ref{fig:MaxwellDT} shows a curve for the liquid, and a curve for 
%the solid both at a fixed temperature. Since the pressure of the liquid 
%is the same as the pressure of the solid at the point of transistion, the 
%density of the liquid and the density of the solid at the point of freezing 
%can be found from identifying the inverse densties at which the tangents 
%to each curve have the same slope.%
Two such curves are shown in Figure \ref{fig:MaxwellDT}, one for the liquid 
and one for the solid at a fixed temperature. Since the pressure of the 
liquid is the same as the pressure of the solid at the point of transistion, 
the density of the liquid and the density of the solid at the point of 
freezing can be found from identifying the inverse densties at which the 
tangents to each curve have the same slope. This process is called 
Maxwell's Double Tangent Construction.

\begin{figure}[h!]
    \centering
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig1.pdf}
     % FIXME\includegraphics[height=7cm]{plot-pressure_Fig1.pdf}
    \caption{Identifying densities at the solid-liquid transition using 
    Maxwell's Double Tangent Construction}
    \label{fig:MaxwellDT}
  \end{figure}

@@

%\begin{displaymath}{F=-PV+G}\end{displaymath} 
\begin{displaymath}{\frac{F}{N}=-P\frac{V}{N}+\frac{G}{N}}\end{displaymath} 
\begin{displaymath}{f=-P\frac{1}{n}+g}\end{displaymath}
This equation has the form of a line:
\begin{displaymath}{y=mx+y_o}\end{displaymath}
%\begin{displaymath}{G_1=G_2}\end{displaymath} 
%\begin{displaymath}{F_1+PV_1=F_2+PV_2}\end{displaymath} 
%\begin{displaymath}{\frac{F_1}{N}+P\frac{V_1}{N}=\frac{F_2}{N}+P\frac{V_2}{N}}\end{displaymath} 
and describes the red line shown in Figure \ref{fig:MaxwellDT} with the 
pressure given by the negative of the slope of the line, and the Gibb's 
free energy per atom given by the y-intercept. At the point of phase transition:
\begin{displaymath}{g_L=g_S}\end{displaymath} 
\begin{displaymath}{f_L+P\frac{1}{n_L}=f_S+P\frac{1}{n_S}}\end{displaymath}

The pressure at the transistion point can be obtained from the point of 
intersection that occurs when the Gibbs free energy per atom is plotted 
against the pressure for both the liquid and the solid at fixed temperature, 
as shown in Figure \ref{fig:GibbsvsP}. These curves are constructed from 
the slope and cooresponding y-intercept data collected from lines tangent 
along the curves on the the Helmholtz free energy per atom verses inverse 
density plots for both the liquid and the solid. 
\begin{figure}[h!]
    \centering
     \includegraphics[height=8cm]{MaxwellDTC-Fig2.pdf}
     %\includegraphics[height=7cm]{plot-pressure_Fig2.pdf}
    \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
    \label{fig:GibbsvsP}
  \end{figure}
%CURVE FITTING! - put in methods/analysis

@@

\begin{figure}[h!]
    \centering
    \includegraphics[height=8cm]{MaxwellDTC-Fig3.pdf}
    % \includegraphics[height=7cm]{plot-pressure_Fig3.pdf}
    \caption{The liquid and solid densities at the point of transistion are 
    found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
    \label{fig:Pvsinvn}
  \end{figure}

A Temperature-Density (T-n) phase diagram, such as that in Figure \ref{fig:T-n_Diagram}, 
can be generated by repeating this procedure for various temperatures. 
%The lowest density the crystal exhibts and the highest density the fluid 
%exhibits at the phase transistion can then be identified for each temperature.

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{T-n_Diagram.png}
    \caption{Conceptual Temperature-Density Phase Diagram used to identify 
    the number densities of the liquid and the solid during a phase 
    transistion. The gray region is where the liquid and the solid coexist, 
    and are therefore not pure states.}
    \label{fig:T-n_Diagram}
  \end{figure} 
%One way of implementing this practically is to determine what the Helmholtz 
%free energy would be for both the liquid and solid phases at a particular 
%temperature and density. Whichever phase has the lower Helmholtz free 
%energy will be the phase that the substance will be in.

@@

%The reason this works is because cDFT satisfies the conditions of the 
%variational principle, the same mathematical principle used to determine 
%that the shortest distance between two points is a straight line by 
%varying the path between the two points until the distance is minimized.
